\chapter{MANIFOLDS}

\textit{July 20th.}

The primary reference for this course will be Loring Tu's \emph{An Introduction to Manifolds}. Prerequisites include Analysis of Several Variables, Linear Algebra, and Topology. The mid-term will account for 35\% of the final grade, quizzes and student lectures will account for 15\%, and the final exam will account for 50\%. 

Simply put, a manifold is any object that is locally Euclidean.

\section{Calculus on $\R^{n}$}
A point $p$ or element in $\R^{n}$ is associated with an $n$-tuple of coordinates $x^{1},x^{2},\ldots,x^{n}$. Let $U$ be an open neighbourhood of $p$ in $\R^{n}$. For a non-negative integer $k$, we say that a function $f:U\to\R$ is \eax{$C^{k}$-continuous} at $p \in U$ if all partial derivatives of $f$ of order $\leq k$ exist and are continuous at $p$. $f$ is said to be $C^{\infty}$ or \eax{smooth} if $f$ is $C^{k}$ for all $k \geq 0$. $f$ is $C^{k}$ (or $C^{\infty}$)on $U$ if it is $C^{k}$ (or $C^{\infty}$) at every point in $U$.

If $f$ happens to be a vector-valued function, i.e. $f:U\to\R^{m}$, we say $f = (f^{1},f^{2},\ldots,f^{m})$ is $C^{k}$ (or $C^{\infty}$) at $p$ if each component function $f^{i}:U\to\R$ is $C^{k}$ (or $C^{\infty}$) at $p$. $f$ is $C^{k}$ (or $C^{\infty}$) on $U$ if it is $C^{k}$ (or $C^{\infty}$) at every point in $U$.

\begin{definition}
    Let $R$ be a commutative ring with unity. We term $A$ to be an $R$-\eax{algebra} if $A$ is a commutative ring with unity and an $R$-module, and for $r \in R$ and $a,b \in A$, we have $r(ab) = (ra)b = a(rb)$.
\end{definition}
In this course, we will be primarily be concerned with the case when $R$ is a field, and $A$ is an $R$-vector space.

For a subset $U \subseteq \R^{n}$, we denote $C^{k}(U)$ (or $C^{\infty}(U)$) as the set of all $C^{k}$ (or $C^{\infty}$) functions from $U$ to $\R$. Note that this is a commutative ring with unity. It is also a vector space over $\R$. Together, $C^{k}(U)$ (or $C^{\infty}(U)$) is a (commutative) algebra over $\R$. Moreover, we have the descending chain of subalgebras
\begin{align}
    C^{0}(U) \supseteq C^{1}(U) \supseteq C^{2}(U) \supseteq \cdots C^{k}(U) \supseteq C^{k+1}(U) \supseteq \cdots
\end{align}
with
\begin{align}
    \bigcap_{k \geq 0} C^{k}(U) = C^{\infty}(U).
\end{align}

$f:U \to \R$ is termed \eax{real analytic} at $p \in U$ if there exists a neighbourhood $V \subseteq U$ of $p$ such that
\begin{align}
    f(x) = f(p) + \sum_{i=1}^{n} \frac{\partial f}{\partial x^{i}}(p)(x^{i}-p^{i}) + \cdots + \frac{1}{k!} \sum_{1 \leq i_{1},\ldots,i_{k} \leq n} \frac{\partial^{k} f}{\partial x^{i_{1}} \cdots \partial x^{i_{k}}}(p)(x^{i_{1}}-p^{i_{1}}) \cdots (x^{i_{k}}-p^{i_{k}}) + \cdots
\end{align}
for all $x \in V$. $f$ is said to be real analytic on $U$ if it is real analytic at every point in $U$. We denote the set of all real analytic functions on $U$ as $C^{\omega}(U)$. Note that $C^{\omega}(U) \subseteq C^{\infty}(U)$.

\begin{theorem}[\eax{Taylor's theorem}]
    Let $p \in U \subseteq \R^{n}$, and $f:U \to \R$ be a $C^{\infty}$ function. Further assume that for all $x \in U$, the line segment from $p$ to $x$ is contained in $U$ ($U$ in this case is said to be \eax{star-shaped} with respect to $p$). Then there exist $g_{1},\ldots,g_{n} \in C^{\infty}(U)$ such that, for all $x \in U$,
    \begin{align}
        f(x) = f(p) + \sum_{i=1}^{n} g_{i}(x)(x^{i}-p^{i}),
    \end{align}
    with
    \begin{align}
        g_{i}(p) = \frac{\partial f}{\partial x^{i}}(p), \quad 1 \leq i \leq n.
    \end{align}
\end{theorem}
\begin{proof}
    Let $x \in U$. Then the line segment $\{p+t(x-p) \mid t \in [0,1]\}$ is contained in $U$. Composing $f$ with the map $t \mapsto p+t(x-p)$, we obtain
    \begin{align}
        \frac{df}{dt}(p+t(x-p)) = \sum_{i=1}^{n} \frac{\partial f}{\partial x^{i}}(p+t(x-p))(x^{i}-p^{i}).
    \end{align}
    Integrating from $0$ to $1$, we get
    \begin{align}
        f(x) - f(p) = \int_{0}^{1} \frac{df}{dt}(p+t(x-p)) dt = \sum_{i=1}^{n} (x^{i}-p^{i}) \int_{0}^{1} \frac{\partial f}{\partial x^{i}}(p+t(x-p)) dt = \sum_{i=1}^{n} (x^{i}-p^{i}) g_{i}(x).
    \end{align}
    Clearly, because $f$ is $C^{\infty}$, each $g_{i}$ is also $C^{\infty}$. Further, we have $g_{i}(p) = \int_{0}^{1} \frac{\partial f}{\partial x^{i}}(p) dt = \frac{\partial f}{\partial x^{i}}(p)$.
\end{proof}

Let $p = (p^{1},\ldots,p^{n}), v \in \R^{n}$, and $f:U \to \R$ be $C^{\infty}$ at $p$. The \eax{directional derivative} of$ f$ at $p$ in the direction of $v$ is defined as
\begin{align}
    \lim_{t \to 0} \frac{f(p+tv)-f(p)}{t} = \frac{d}{dt}f(p+tv)\Big|_{t=0} = \sum_{i=1}^{n} \frac{df}{dx^{i}}(p)v^{i}.
\end{align}
From here, we obtain the operator
\begin{align}
    D_{v} = \sum_{i=1}^{n} v^{i} \frac{\partial}{\partial x^{i}}\Big|_{p}
\end{align}
Now consider all pairs of the form $(f,U)$ where $p \in U \subseteq \R^{n}$ is an open neighbourhood of $p$ and $f:U \to \R$ is $C^{\infty}$. Define an equivalence relation $\sim$ on this set by declaring $(f,U) \sim (g,V)$ if and only if there exists an open neighbourhood $W \subseteq U \cap V$ of $p$ such that $f|_{W} = g|_{W}$. The equivalence class of $(f,U)$ is denoted by $[(f,U)]$, and the set of all such equivalence classes is denoted by $C^{\infty}_{p}(\R^{n})$. Such an equivalence class $[(f,U)]$ is called the \eax{germ} of $f$ at $p$. We can give $C^{\infty}_{p}(\R^{n})$ the structure of a commutative ring with unity by defining
\begin{align}
    [(f,U)] + [(g,V)] = [(f+g,U \cap V)], \quad [(f,U)] \cdot [(g,V)] = [(fg,U \cap V)], \quad 1 = [(1,\R^{n})].
\end{align}
It can also be given the structure of a vector space over $\R$ by defining
\begin{align}
    \lambda [(f,U)] = [(\lambda f,U)], \quad \lambda \in \R.
\end{align}
Thus, $C^{\infty}_{p}(\R^{n})$ is an $\R$-algebra. Now, given a germ $\alpha = [(f,U)] \in C^{\infty}_{p}(\R^{n})$ and a vector $v \in \R^{n}$, we define
\begin{align}
    D_{v}\alpha = \sum_{i=1}^{n} v^{i} \frac{\partial f}{\partial x^{i}}\Big|_{p} = v \cdot \nabla f(p).
\end{align}
Note that the choice of $f$ is immaterial, since if $[(f,U)] = [(g,V)]$, then there exists an open neighbourhood $W \subseteq U \cap V$ of $p$ such that $f|_{W} = g|_{W}$, and hence $\nabla f(p) = \nabla g(p)$.
One can infer that $D_{\lambda v}(f) = \lambda D_{v}(f) = D_{v}(\lambda f)$ and $D_{v}(f+g) = D_{v}(f) + D_{v}(g)$ for all $\lambda \in \R$ and $f,g \in C^{\infty}_{p}(\R^{n})$. Thus, $D_{v}$ is an $\R$-linear map from $C^{\infty}_{p}(\R^{n})$ to $\R$ (hereforth, we shall use $f$ to also denote the germ $[(f,U)]$ of $f$ at $p$). Also,
\begin{align}
    D_{v}(fg) = \sum_{i=1}^{n} v^{i} \frac{\partial (fg)}{\partial x^{i}}(p) = \sum_{i=1}^{n} v^{i} \left( f(p) \frac{\partial g}{\partial x^{i}}(p) + g(p) \frac{\partial f}{\partial x^{i}}(p) \right) = f(p) D_{v}(g) + g(p) D_{v}(f).
\end{align} 
Such a property makes $D_{v}$ what we call a derivation.
\begin{definition}
    Any $\R$-linear map $T:C^{\infty}_{p}(\R^{n}) \to \R$ satisfying the Leibniz rule $T(fg) = f(p)T(g) + g(p)T(f)$ for all $f,g \in C^{\infty}_{p}(\R^{n})$ is called a \eax{derivation} at $p$.
\end{definition}
One may show that if $T$ is a derivation at $p$, then $T(1) = 0$. 

The set of all derivations at $p$ is denoted by $\mf{D}_{p}(\R^{n})$. It is easy to see that $\mf{D}_{p}(\R^{n})$ is a vector space over $\R$. Now consider the map
\begin{align}
    \theta:\R^{n} \to \mf{D}_{p}(\R^{n}), \quad v \mapsto D_{v}.
\end{align}
This $\theta$ is an $\R$-linear map (verify). We claim that $\theta$ is an isomorphism of real vector spaces. To this end, let $v \in \R^{n}$ such that $D_{v} = 0 \in \mf{D}_{p}(\R^{n})$. Then $D_{v}f = 0$ for all $f \in C^{\infty}_{p}(\R^{n})$. In particular, for the coordinate functions $x^{i}$, we have $D_{v}x^{i} = v_{i} = 0$ for all $1 \leq i \leq n$. Thus, $v = 0$, and hence $\theta$ is injective. To show surjectivity, we appeal to Taylor's theorem; let $D \in \mf{D}_{p}(\R^{n})$. We claim that $D = D_{v}$ where $v = (D(x^{1}),\ldots,D(x^{n}))$. Indeed, for any $f \in C^{\infty}_{p}(\R^{n})$, we have
\begin{align}
    D(f) &= D(f(p)) + D\left( \sum_{i=1}^{n} (x^{i}-p^{i}) g_{i}(x) \right) = 0 + \sum_{i=1}^{n} \left( (x^{i}-p^{i})(p) Dg_{i}(x) + D(x^{i}-p^{i})g_{i}(p) \right) \notag \\
    &= \sum_{i=1}^{n} (Dx^{i})g_{i}(p) = \sum_{i=1}^{n} \frac{\partial f}{\partial x^{i}}(p) Dx^{i} = D_{v}(f).
\end{align}
Hence, $\theta$ is surjective, and we have the desired isomorphism $\mf{D}_{p}(\R^{n}) \cong \R^{n}$.

Note that a germ $f \in C^{\infty}_{p}(\R^{n})$ has a valid value at $p$, but no other point. Hence, we have the evaluation map 
\begin{align}
    \ev_p: C^{\infty}_{p}(\R^{n}) \to \R, \quad [(f,U)] \mapsto f(p).
\end{align}
The kernel here is $\{[(f,U)] \in C^{\infty}_{p}(\R^{n}) \mid f(p) = 0\}$, which is a ideal in $C^{\infty}_{p}(\R^{n})$. For any other element $[(f,U)] \in C^{\infty}_{p}(\R^{n})$ with $f(p) \neq 0$, a valid inverse is given by $[(f^{-1},U\setminus\{f=0\})]$. This makes $C^{\infty}_{p}(\R^{n})$ a local ring.
\\

\section{Multilinear Algebra}

\textit{July 22nd.}

Hereforth, $V$ shall denote a finite-dimensional real vector space, and $V^{V}$ shall denote its \eax{dual space} $\Hom_{\R}(V,\R)$. If $e_{1},\ldots,e_{n}$ are a basis for $V$, then the \eax{dual basis} of $V^{V}$ is denoted by $\alpha^{1},\ldots,\alpha^{n} \in V^{V}$, and is defined by $\alpha^{i}(e_{j}) = \delta_{i,j}$ for all $1 \leq i,j \leq n$. One may verify that the dual basis $\{\alpha^{1},\ldots,\alpha^{n}\}$, induced by the basis $\{e_{1},\ldots,e_{n}\}$, is indeed a basis for $V^{V}$. $S_{k}$ shall denote the permutation group on $k$ symbols, satisfying $\abs{S_{k}} = k!$. For any $\sigma \in S_{k}$, the sign of $\sigma$ is denoted by $\sgn(\sigma)$, and is defined as $+1$ if $\sigma$ is an even permutation, and $-1$ if $\sigma$ is an odd permutation. 

With notation out of the way, let $k \geq 1$ be an integer. We shall consider the $k$-tuple $V^{k} = V \times \cdots \times V$ ($k$ times).

\begin{definition}
    A mapping $f:V^{k} \to \R$ is said to be \eax{$k$-linear} if $f$ is linear in each of its $k$ arguments.
\end{definition}
A special case is when $k = 2$, in which case $f$ is called a \eax{bilinear form}. Such an $f:V \times V \to \R$ satisfies
\begin{align}
    f(\lambda v + \mu w, u) = \lambda f(v,u) + \mu f(w,u), \quad f(u,\lambda v + \mu w) = \lambda f(u,v) + \mu f(u,w).
\end{align}
A common example of a bilinear form is any valid inner product $\ip{\cdot,\cdot}:V \times V \to \R$ on $V$. Another example is the determinant $\det:\R^{n} \to \R$, where the vectors are interpreted as the columns of a matrix. We can also write
\begin{align}
    L_{k}(V) = \{f:V^{k} \to \R \mid f \text{ is $k$-linear}\}.
\end{align}
It is easy to see that $L_{k}(V)$ is an $\R$-vector space.

\begin{definition}
    A $k$-linear map $f:V^{k} \to \R$ is called a \eax{symmetric $k$-linear form} if for all $\sigma \in S_{k}$,
    \begin{align}
        f(v_{1},\ldots,v_{k}) = f(v_{\sigma(1)},\ldots,v_{\sigma(k)}).
    \end{align}
    That is, $f$ is invariant under any permutation of its arguments. Moreover, $f$ is called an \eax{alternating $k$-linear form} if for all $\sigma \in S_{k}$,
    \begin{align}
        f(v_{1},\ldots,v_{k}) = \sgn(\sigma)f(v_{\sigma(1)},\ldots,v_{\sigma(k)}).
    \end{align}
\end{definition}
One can see that the standard inner product on $\R^{n}$ is a symmetric bilinear form, and the determinant is an alternating $n$-linear form. Related to these definitions, we can immediately define the following subspaces of $L_{k}(V)$:
\begin{align}
    S_{k}(V) = \{f \in L_{k}(V) \mid f \text{ is symmetric}\} \subseteq L_{k}(V), \quad A_{k}(V) = \{f \in L_{k}(V) \mid f \text{ is alternating}\} \subseteq L_{k}(V).
\end{align}
Currently, our main interest of study is $A_{k}(V)$, the space of alternating $k$-linear forms on $V$. Elements of $A_{k}(V)$ are termed \eax{alternating $k$-tensors}. Similarly, those of $S_{k}(V)$ and $L_{k}(V)$ are called \eax{symmetric $k$-tensors} and \eax{$k$-tensors}, respectively. For $k = 0$, we define $L_{0}(V) = S_{0}(V) = A_{0}(V) = \R$. For $k = 1$, we have $L_{1}(V) = S_{1}(V) = A_{1}(V) = V^{V}$. We can define the (left) action of $S_{k}$ on $L_{k}(V)$ by
\begin{align}
    (\sigma \cdot f)(v_{1},\ldots,v_{k}) = f(v_{\sigma(1)},\ldots,v_{\sigma(k)}).
\end{align}
Thus, $f$ is symmetric if $\sigma \cdot f = f$ for all $\sigma \in S_{k}$, and $f$ is alternating if $\sigma \cdot f = \sgn(\sigma)f$ for all $\sigma \in S_{k}$.

Given an $f \in L_{k}(V)$, we can define the \eax{symmetrization} $Sf \in S_{k}(V)$, of $f$, as
\begin{align}
    Sf = \sum_{\sigma \in S_{k}} \sigma \cdot f.
\end{align}
One may, indeed, check that $Sf$ is symmetric. Similarly, we can define the \eax{alternation} $Af \in A_{k}(V)$, of $f$, as
\begin{align}
    Af = \sum_{\sigma \in S_{k}} \sgn(\sigma) \sigma \cdot f.
\end{align}

\subsection{The Tensor and Wedge Products}

\noindent \textit{July 24th.}

For any $f \in L_{k}(V)$ and $g \in L_{l}(V)$, we can define the \eax{tensor product} $f \otimes g \in L_{k+l}(V)$ as
\begin{align}
    (f \otimes g)(v_{1},\ldots,v_{k+l}) = f(v_{1},\ldots,v_{k})g(v_{k+1},\ldots,v_{k+l}).
\end{align}
Verify that $f \otimes g$ is indeed $(k+l)$-linear. We are interested in the tensor product of alternating tensors, so let $f \in A_{k}(V)$ and $g \in A_{l}(V)$. Then, we can define the \eax{wedge product} $f \wedge g \in A_{k+l}(V)$ as
\begin{align}
    f \wedge g = \frac{1}{k!l!} A(f \otimes g) = \frac{1}{k!l!} \sum_{\sigma \in S_{k+l}} \sgn(\sigma) \sigma \cdot (f \otimes g).
\end{align}
The associativity of the tensor product holds easily: for $f \in L_{k}(V)$, $g \in L_{l}(V)$, and $h \in L_{m}(V)$,
\begin{align}
    (f \otimes g) \otimes h = f \otimes (g \otimes h).
\end{align}

\begin{proposition}
    For $f \in A_{k}(V)$ and $g \in A_{l}(V)$, we have
    \begin{align}
        f \wedge g = (-1)^{kl} g \wedge f.
    \end{align}
\end{proposition}
\begin{proof}
    We shall show that $A(f \otimes g) = (-1)^{kl} A(g \otimes f)$. To this end, fix the permutation $\tau \in S_{k+l}$ defined by
    \begin{align}
        \tau = \begin{pmatrix}
            1 & 2 & \cdots & l & l+1 & l+2 & \cdots & l+k \\
            k+1 & k+2 & \cdots & k+l & 1 & 2 & \cdots & k
        \end{pmatrix}.
    \end{align}
    We have
    \begin{align}
        A(f \otimes g)(v_{1},\ldots,v_{k+l}) &= \sum_{\sigma \in S_{k+l}} \sgn(\sigma) f(v_{\sigma(1)},\ldots,v_{\sigma(k)}) g(v_{\sigma(k+1)},\ldots,v_{\sigma(k+l)}) \notag \\
        &= \sum_{\sigma \in S_{k+l}} \sgn(\sigma) f(v_{\sigma\tau(l+1)},\ldots,v_{\sigma\tau(l+k)}) g(v_{\sigma\tau(1)},\ldots,v_{\sigma\tau(l)}) \notag \\
        &= \sgn(\tau) \sum_{\sigma \in S_{k+l}} \sgn(\sigma\tau) f(v_{\sigma\tau(l+1)},\ldots,v_{\sigma\tau(l+k)}) g(v_{\sigma\tau(1)},\ldots,v_{\sigma\tau(l)}) \notag \\
        &= \sgn(\tau)A(g \otimes f)(v_{1},\ldots,v_{k+l}).
    \end{align}
    With $\sgn(\tau) = (-1)^{kl}$, the proof is complete.
\end{proof}

\begin{corollary}
    For $f \in A_{k}(V)$, we have $f \wedge f = 0$ if $k$ is odd.
\end{corollary}
\begin{proof}
    Easy since $f \wedge f = (-1)^{k^{2}} f \wedge f = -f \wedge f$ if $k$ is odd, so $f \wedge f = 0$.
\end{proof}

\begin{lemma}
    For $f \in L_{k}(V)$ and $g \in L_{l}(V)$, we have
    \begin{align}
        A(Af \otimes g) = k!A(f \otimes g), \quad A(f \otimes Ag) = l!A(f \otimes g).
    \end{align}
\end{lemma}

\begin{proof}
    We shall prove the first equality; the second follows similarly. We have
    \begin{align}
        A(Af \otimes g) &= \sum_{\sigma \in S_{k+l}} \sgn(\sigma) \sigma \cdot \left( \sum_{\tau \in S_{k}} \sgn(\tau) (\tau \cdot f)  \otimes g \right) = \sum_{\sigma \in S_{k+l}} \sgn(\sigma) \sigma \cdot \left( \sum_{\tau \in S_{k}} \sgn(\tau) \tau \cdot(f \otimes g) \right) \notag \\
        &= \sum_{\tau \in S_{k}} \sum_{\sigma \in S_{k+l}} \sgn(\sigma \tau) \sigma \tau \cdot (f \otimes g) = k! \sum_{\alpha \in S_{k+l}} \sgn(\alpha) \alpha \cdot (f \otimes g) = k! A(f \otimes g).
    \end{align}
\end{proof}

\begin{proposition}
    The wedge product is associative, i.e. for $f \in A_{k}(V)$, $g \in A_{l}(V)$, and $h \in A_{m}(V)$, we have
    \begin{align}
        (f \wedge g) \wedge h = f \wedge (g \wedge h).
    \end{align}
\end{proposition}
\begin{proof}
    We have
    \begin{align}
        (f \wedge g) \wedge h &= \frac{1}{(k+l)!m!} A((f \wedge g) \otimes h) = \frac{1}{(k+l)!m!k!l!} A(A(f \otimes g) \otimes h) = \frac{(k+l)!}{(k+l)!l!k!m!}A((f \otimes g) \otimes h) \notag \\
        &= \frac{1}{k!l!m!} A(f \otimes g \otimes h) = f \wedge (g \wedge h).
    \end{align}
\end{proof}

\begin{corollary}
    $f \wedge g \wedge h = \frac{1}{k!l!m!} A(f \otimes g \otimes h)$ for $f \in A_{k}(V)$, $g \in A_{l}(V)$, and $h \in A_{m}(V)$. Inductively,
    \begin{align}
        f_{1} \wedge \cdots \wedge f_{k} = \frac{1}{k_{1}!\cdots k_{k}!} A(f_{1} \otimes \cdots \otimes f_{k}),
    \end{align}
    where $f_{i} \in A_{k_{i}}(V)$ for $1 \leq i \leq k$.
\end{corollary}

\begin{proposition}
    Let $\alpha^{1},\ldots,\alpha^{n} \in V^{V} = A_{1}(V) = L_{1}(V)$, and $v_{1},\ldots,v_{k} \in V$. Then
    \begin{align}
        (\alpha^{1} \wedge \cdots \wedge \alpha^{k})(v_{1},\ldots,v_{k}) = \det(\alpha^{i}(v_{j}))_{1 \leq i,j \leq k}.
    \end{align}
\end{proposition}

\begin{proof}
    We have
    \begin{align}
        (\alpha^{1} \wedge \cdots \wedge \alpha^{k})(v_{1},\ldots,v_{k}) &= \sum_{\sigma \in S_{k}} \sgn(\sigma)(\alpha^{1} \otimes \cdots \otimes \alpha^{k})(v_{\sigma(1)},\ldots,v_{\sigma(k)}) \notag \\
        &= \sum_{\sigma \in S_{k}} \sgn(\sigma) \alpha^{1}(v_{\sigma(1)})\alpha^{2}(v{\sigma(2)}) \cdots \alpha^{k}(v_{\sigma(k)}) = \det(\alpha^{i}(v_{j}))_{1 \leq i,j \leq k}.
    \end{align}
\end{proof}

Let $V$ be an $\R$-vector space with basis $e_{1},\ldots,e_{n}$, and $V^{V}$ be its dual space with dual basis $\alpha^{1},\ldots,\alpha^{n}$. Moreover, let $I$ be a multi-index of length $k$, i.e. $I = (i_{1},\ldots,i_{k})$ with $1 \leq i_{1},\ldots,i_{k} \leq n$. We define $e_{I} \defeq (e_{i_{1}},\ldots,e_{i_{k}}) \in V^{k}$ and $\alpha^{I} \defeq \alpha^{i_{1}} \wedge \cdots \wedge \alpha^{i_{k}} \in A_{k}(V)$. Then, a $k$-linear form $f \in L_{k}(V)$ is completely determined by its values on the $e_{I}$, and an alternating $k$-linear form $f \in A_{k}(V)$ is completely determined by its values on the $e_{I}$ with $i_{1} < i_{2} < \cdots < i_{k}$. 