\chapter{REFINING THE NOTION OF PROBABILITY}

\textit{July 21st.}

Consider the random walk on the line; a drunk ant sits at the origin and takes steps of length $1$ to the left or right with equal probability. We wish to obtain the probability that after an even $n$ steps, the ant returns to the origin. Of course, the natural approach is to frame this problem in a discrete probability space.

\begin{definition}
    A \eax{discrete probability space} is a pair $(\Omega,q)$ where $\Omega$ is a non-empty finite or countably infinite set called the sample space, and $q:\Omega\to[0,1]$ is a function satisfying $\sum_{\omega\in\Omega}q(\omega)=1$. Any subset of $\Omega$ is termed an event, and the probability of an event $A \subseteq \Omega$ is defined as $\Prob(A) = \sum_{\omega\in A}q(\omega)$.
\end{definition}

For the problem above, we can take the sample space to be
\begin{align}
    \Omega = \{\bv{x}=(x_{0}=0,x_{1},x_{2},\ldots,x_{n}) \in \Z^{n+1} \mid x_{i+1}-x_{i} \in \{-1,1\} \text{ for all } 0 \leq i < n\},
\end{align}
and the mapping $q$ to be $q(\bv{x}) = 2^{-n}$ for all $\bv{x} \in \Omega$. Then, the event that the ant returns to the origin after $n$ steps is given by $A = \{\bv{x} \in \Omega \mid x_{n} = 0\}$, and we have
\begin{align}
    \Prob(A) = \sum_{\bv{x} \in A} q(\bv{x}) = 2^{-n} \abs{A} = 2^{-n} \binom{n}{n/2}.
\end{align}
Reframing the problem, let us consider a random walk on the line starting at the origin, and continuing indefinitely. In this case, the sample space is given by
\begin{align}
    \Omega = \{\bv{x}=(x_{0}=0,x_{1},x_{2},\ldots) \in \Z^{\N} \mid x_{i+1}-x_{i} \in \{-1,1\} \text{ for all } i \in \N\}.
\end{align}
This $\Omega$ is uncountably infinite. In such a case, we cannot define a probability measure by assigning a positive probability to each element of $\Omega$. As another example of difficulty in dealing with uncountable sample spaces, consider the problem of choosing a real number uniformly at random from the interval $[0,1]$. If we attempt to assign a positive probability to each element of $[0,1]$, we would have to assign a probability of $0$ to each element, since there are uncountably many elements in $[0,1]$. However, this would lead to a total probability of $0$, which is not valid. Of course, we know that the probability of choosing a specific element or a countable set of elements from $[0,1]$ is $0$, but there then arises the question of how to assign a probability to an uncountable set of elements, such as the Cantor set. We can work in a slightly different viewpoint.

\begin{definition}
    A probability space is a tuple $(\Omega,\cF,\Prob)$ where $\Omega$ is a non-empty set called the sample space, $\cF$ is the collection of all subsets of $\Omega$ ($\sim 2^{\Omega}$), and $\Prob:\cF\to[0,1]$ is a function satisfying $\Prob(\Omega)=1$ and countable additivity, that is, for any countable collection of pairwise disjoint sets $A_{1},A_{2},\ldots \in \cF$, we have
    \begin{align}
        \Prob\left(\bigcup_{i=1}^{\infty}A_{i}\right) = \sum_{i=1}^{\infty}\Prob(A_{i}).
    \end{align}
\end{definition}

Using this definition, we look to define a $\Prob$ on subsets of $[0,1]$ that satisfies the above properties, along with the property that $\Prob([a,b]) = b-a$. Thus, we have $\Prob([0,1]) = 1$. If we have two disjoint intervals $I_{1},I_{2} \subseteq [0,1]$, then $\Prob(I_{1}\cup I_{2}) = \Prob(I_{1}) + \Prob(I_{2})$. This is a natural property to expect from a probability measure. What about a general set $S \subseteq [0,1]$? One can observe that if we have a sequence of intervals $I_{1},I_{2},\ldots$ such that $S \subseteq \bigcup_{k=1}^{\infty}I_{k}$, then we would have $\Prob(S) \leq \sum_{k=1}^{\infty}\Prob(I_{k})$. This is because the probability of $S$ should be less than or equal to the probability of any set that contains $S$. Thus, we can define
\begin{align}
    \Prob_{\ast}(S) = \inf \left\{ \sum_{k=1}^{\infty}\abs{I_{k}} \, \Big| \, S \subseteq \bigcup_{k=1}^{\infty}I_{k}, \; I_{k} \subseteq [0,1] \right\}.
\end{align}
We have not verified that this is a valid definition of a probability measure, hence the subscript $\ast$. Via this, let us try computing $\Prob_{\ast}(\Q \cap [0,1])$. Enumerate the rationals in $[0,1]$ as $\{q_{1},q_{2},\ldots\}$. For any $\epsilon > 0$, we can choose intervals $I_{k} = [q_{k}-\epsilon/2^{k+1},q_{k}+\epsilon/2^{k+1}] \cap [0,1]$ for each $k \in \N$. Note that $\bigcup_{k=1}^{\infty}I_{k}$ contains all the rationals in $[0,1]$, and $\sum_{k=1} \abs{I_{k}} \leq \sum_{k=1}^{\infty}\epsilon/2^{k} = \epsilon$. Since $\epsilon$ was arbitrary, we have $\Prob_{\ast}(\Q \cap [0,1]) = 0$. This is consistent with our intuition that the probability of choosing a rational number from $[0,1]$ is $0$. So, $\Prob_{\ast}$ seems to be a reasonable definition of a probability measure on $([0,1],2^{[0,1]})$. However, we will show later that there exists a set $\cA \subseteq [0,1]$ such that $\Prob_{\ast}(\cA) = \Prob_{\ast}([0,1]\setminus \cA) = 1$, contradicting the property that $\Prob_{\ast}([0,1]) = 1$. We state the following theorem without proof.
\begin{theorem}
    There is no probability measure $\Prob$ on $([0,1],2^{[0,1]})$ such that $\Prob([a,b]) = b-a$ for all $[a,b] \subseteq [0,1]$.
\end{theorem}
If we remove the requirement that the probability measure of a subinterval $[a,b]$ is equal to its length, then we can define a probability measure on $([0,1],2^{[0,1]})$. For example, we can define $\Prob(S) = 1$ if $1/2 \in S$ and $\Prob(S) = 0$ otherwise. This is a valid probability measure, but it is not very useful. Keeping the requirement that the probability measure of a subinterval $[a,b]$ is equal to its length, we can instead restrict the collection of subsets of $[0,1]$ on which we define the probability measure. 

\begin{definition}
    A \eax{probability space} is a tuple $(\Omega,\cF,\Prob)$ where
    \begin{itemize}
        \item $\Omega$ is a non-empty set called the sample space,
        \item $\cF$ is a collection of subsets of $\Omega$ satisfying the following properties:
        \begin{itemize}
            \item $\Omega \in \cF$,
            \item if $A \in \cF$, then $\Omega \setminus A \in \cF$,
            \item if $A_{1},A_{2},\ldots \in \cF$, then $\bigcup_{i=1}^{\infty}A_{i} \in \cF$,
        \end{itemize}
        \item $\Prob:\cF\to[0,1]$ is a function satisfying:
        \begin{itemize}
            \item $\Prob(\Omega) = 1$,
            \item if $A_{1},A_{2},\ldots \in \cF$ are pairwise disjoint, then $\Prob\left(\bigcup_{i=1}^{\infty}A_{i}\right) = \sum_{i=1}^{\infty}\Prob(A_{i})$.
        \end{itemize}
    \end{itemize}
\end{definition}
Such an $\cF$ is termd a \eax{$\sigma$-algebra} on $\Omega$, and such a $\Prob$ is termed a \eax{probability measure} on $(\Omega,\cF)$.

\section{$\sigma$-algebras}

\begin{example}
    Given $\Omega$, trivial possible choices for $\cF$ are $\{\emptyset,\Omega\}$ and $2^{\Omega}$. A more interesting example is the following: suppose $\Omega$ is uncountable. Then $\cF = \{A \subseteq \Omega \mid A \text{ is countable or } \Omega \setminus A \text{ is countable}\}$ is a $\sigma$-algebra on $\Omega$.
\end{example}

Let $S \subseteq 2^{\Omega}$ be a collection of subsets of $\Omega$. We claim that there exists a notion of the smallest $\sigma$-algebra containing $S$. This is called the $\sigma$-algebra generated by $S$, and is denoted by $\sigma(S)$.

\begin{lemma}
    For any collection $S \subseteq 2^{\Omega}$, there exists a unique smallest $\sigma$-algebra containing $S$ satisfying
    \begin{align}
        \sigma(S) = \bigcap \{\cF \subseteq 2^{\Omega} \mid \cF \text{ is a $\sigma$-algebra and } S \subseteq \cF\}.
    \end{align}
\end{lemma}
\begin{proof}
    It is easy to check that $\sigma(S)$ is indeed a $\sigma$-algebra containing $S$. If $\cF$ is any $\sigma$-algebra containing $S$, then $\sigma(S) \subseteq \cF$. Thus, $\sigma(S)$ is the unique smallest $\sigma$-algebra containing $S$.
\end{proof}

\begin{definition}
    Let $(X,d)$ be a metric space (or a topological space), and let $\cS \subseteq 2^{X}$ be the colletion of all open sets in $X$. Then, the $\sigma$-algebra $\sigma(\cS)$ is called the \eax{Borel $\sigma$-algebra} on $X$, and its elements are called \eax{Borel set}s.
\end{definition}
We will show that there exists a probability measure on $([0,1],\cB)$, where $\cB$ is the Borel $\sigma$-algebra on $[0,1]$, which extends the notion of length of intervals to Borel sets.

Let us return to the problem of indefinite random walks on the line. Let $\Omega = \{\omega = (\omega_{1},\omega_{2},\ldots) \mid \omega_{i} \in \{0,1\}\}$, where $\omega_{i} = 0$ corresponds to a step to the left and $\omega_{i} = 1$ corresponds to a step to the right. If we let $A = \{0,1\}$, then $\Omega = A^{\N}$; consider the discrete topology on $A$, and the product topology on $\Omega$. If we let
\begin{align}
    \cS = \{U_{1} \times U_{2} \times \cdots \times U_{k} \times A \times A \cdots \mid U_{i} = \{0\} \text{ or } \{1\}\}
\end{align}
then one can show that the $\sigma$-algebra generated by $\cS$ is the Borel $\sigma$-algebra on $\Omega$. We can set $\cF = \sigma(\cS)$, and define a probability measure $\Prob$ on $(\Omega,\cF)$ by setting $\Prob(U_{1} \times U_{2} \times \cdots \times U_{k} \times A \times A \cdots) = 2^{-k}$ for all $U_{i} = \{0\}$ or $\{1\}$.
\\

\textit{July 24th.}
\begin{example}
    Let $(X,d)$ be a separable metric space (that is, there exists a countable dense subset). The Borel $\sigma$-algebra on $X$ is then
    \begin{align}
        \cB(X) = \sigma(\{B(p,r)\}_{p \in X, r >0}).
    \end{align}
    Note that the usual containment is $\sigma(\{B(p,r)\}_{p \in X, r >0}) \subseteq \cB(X)$ for a metric space $X$. So, it suffices to show that if $V \in \tau$, the topology on $X$, then $V \in \sigma(\{B(p,r)\}_{p \in X, r >0})$. Let $Q$ be a countable dense subset of $X$. Then, for any $V \in \tau$, we have
    \begin{align}
        V = \bigcup_{n=1}^{\infty} B(q_{n},r_{n}),
    \end{align}
    where $q_{n} \in Q \cap V$ and $r_{n} > 0$ (where $r_{n}$ is the largest possible radius such that $B(q_{n},r_{n}) \subseteq V$). Since $\sigma(\{B(p,r)\}_{p \in X, r >0})$ is a $\sigma$-algebra, we have $V \in \sigma(\{B(p,r)\}_{p \in X, r >0})$. Thus, we have $\cB(X) = \sigma(\{B(p,r)\}_{p \in X, r >0})$.
\end{example}

\begin{example}
    Let $(\Sigma,\cF,\Prob)$ be a probability space.
    
    \begin{itemize}
        \item $\cF$ is closed under countable unions, countable intersections, differences, and symmetric differences. Countable unions and intersections follow from the definition of a $\sigma$-algebra. If $A,B \in \cF$, then $A \setminus B = A \cap (\Omega \setminus B) \in \cF$. Finally, $A \triangle B = (A \setminus B) \cup (B \setminus A) \in \cF$.

        \item Let $(\Sigma,\cF,\Prob)$ be a probability space, and let $A_{1},A_{2},\ldots \in \cF$. Define
        \begin{align}
            \limsup_{n \to \infty} A_{n} \defeq \{x \in \Omega \mid x \text{ belongs to infinitely many } A_{n}\}
        \end{align}
        and
        \begin{align}
            \liminf_{n \to \infty} A_{n} \defeq \{x \in \Omega \mid x \text{ belongs to all but finitely many } A_{n}\}.
        \end{align}
        Then, both the $\limsup$ and $\liminf$ belong in $\cF$. To see this, simply note that
        \begin{align}
            \limsup_{n \to \infty} A_{n} = \bigcap_{n=1}^{\infty}\bigcup_{k=n}^{\infty}A_{k}, \quad \liminf_{n \to \infty} A_{n} = \bigcup_{n=1}^{\infty}\bigcap_{k=n}^{\infty}A_{k}.
        \end{align}

        \item One may show that $\Prob(\emptyset) = 0$, $\Prob(A) \leq \Prob(B)$ if $A \subseteq B$, $\Prob(A \cup B) = \Prob(A) + \Prob(B) - \Prob(A \cap B)$, and $\Prob(\bigcup A_{n}) \leq \sum \Prob(A_{n})$ for any countable collection of sets $A_{1},A_{2},\ldots \in \cF$.
        
        \item For a nested sequence of sets $A_{1} \subseteq A_{2} \subseteq \cdots$ in $\cF$, let $A = \bigcup_{n=1}^{\infty}A_{n}$ (this is usually denoted by $A_{n} \uparrow A$). Then, $\Prob(A) = \lim_{n \to \infty}\Prob(A_{n})$. To this end, let $B_{n} = A_{n}\setminus A_{n-1}$ for $n > 1$, and $B_{1} = A_{1}$. Then $A = \bigcup_{n=1}^{\infty}B_{n}$. Thus,
        \begin{align}
            \Prob(A) = \sum_{n=1}^{\infty}\Prob(B_{n}) = \lim_{N \to \infty}\sum_{n=1}^{N}\Prob(B_{n}) = \lim_{N \to \infty}\sum_{n=2}^{N} \Prob(A_{n}) - \Prob(A_{n-1}) + \Prob(A_{1}) = \lim_{N \to \infty}\Prob(A_{N}).
        \end{align}
    \end{itemize}    
\end{example}

\subsection{Other Systems of Sets}

The following are other interesting collection of subsets of $\Omega$ besides the Borel $\sigma$-algebra. 

\begin{enumerate}
    \item \eax{$\pi$-system}: a collection of subsets $S \subseteq \Omega$ such that $A,B \in S$ implies $A \cap B \in S$.
    \item \eax{$\lambda$-system}: a collection $\Lambda$ of subsets of $\Omega$ such that $\Omega \in \Lambda$, if $A,B \in \Lambda$ with $A \subseteq B$, then $B \setminus A \in \Lambda$, and if $A_{1},A_{2},\ldots \in \Lambda$, then $\bigcup_{n=1}^{\infty}A_{n} \in \Lambda$.
    \item \eax{algebra}: a collection $\cA$ of subsets of $\Omega$ such that $\Omega \in \cA$, if $A \in \cA$, then $\Omega \setminus A \in \cA$, and if $A,B \in \cA$, then $A \cup B \in \cA$.
\end{enumerate}

We had defined $\omega(S)$, the $\sigma$-algebra generated by a collection of subsets $S \subseteq 2^{\Omega}$. We can similarly define $\pi(S)$, the $\pi$-system generated by $S$, $\lambda(S)$, the $\lambda$-system generated by $S$, and $\cA(S)$, the algebra generated by $S$. 

\begin{theorem}[\eax{SierpiŇĄski-Dynkin $\pi$-$\lambda$ Theorem}]
    Let $S,\cF \subseteq 2^{\Omega}$ be collections of subsets of $\Omega$. Then,
    \begin{enumerate}
        \item $\cF$ is a $\sigma$-algebra if and only if $\cF$ is both a $\pi$-system and a $\lambda$-system.
        \item If $S$ is a $\pi$-system, then $\lambda(S) = \sigma(S)$.
    \end{enumerate}
\end{theorem}
\begin{proof}
    \begin{enumerate}
        \item If $\cF$ is a $\sigma$-algebra, then it is trivially both a $\pi$-system and a $\lambda$-system. Conversely, if $\cF$ is both a $\pi$-system and a $\lambda$-system, we need to verify the axioms of a $\sigma$-algebra. We have $\Omega \in \cF$ since $\cF$ is a $\lambda$-system. If $A \in \cF$, then $\Omega \setminus A \in \cF$ since $\cF$ is a $\lambda$-system. Finally, if $A_{1},A_{2},\ldots \in \cF$, then we can write $B_{n} = \bigcup_{i=1}^{n} A_{i} \in \cF$ since $A_{i}^{c} \in \cF$ and $B_{n}^{c} = \bigcap_{i=1}^{n} A_{i}^{c} \in \cF$ since $\cF$ is a $\pi$-system. Then, $B_{n} \uparrow B = \bigcup_{i=1}^{\infty}A_{i}$, and since $\cF$ is a $\lambda$-system, we have $B \in \cF$. Thus, $\cF$ is a $\sigma$-algebra.
        \item Let $S$ be a $\pi$-system. Since $\lambda(S) \subseteq \sigma(S)$, it suffices to show that $\lambda(S)$ is a $\pi$-system. Fix $A \in S$, and define $\cC_{A} = \{B \in \lambda(S) \mid A \cap B \in \lambda(S)\}$. Note that $S \subseteq \cC_{A}$ because $S$ is a $\pi$-system. We show that $\cC_{A}$ is a $\lambda$-system. We have $\Omega \in \cC_{A}$ since $A \cap \Omega = A \in \lambda(S)$. If $B_{1},B_{2} \in \cC_{A}$ with $B_{1} \subseteq B_{2}$, then $A \cap B_{1} \subseteq A \cap B_{2}$, and since $A \cap B_{2} \in \lambda(S)$, we have $(A \cap B_{2})\setminus(A \cap B_{1}) = A \cap (B_{2}\setminus B_{1}) \in \lambda(S)$, and thus $B_{2}\setminus B_{1} \in \cC_{A}$. Finally, if $B_{1},B_{2},\ldots$ are in $\cC_{A}$, then $A \cap B_{n} \in \lambda(S)$ for all $n$, and thus $\bigcup (A\cap B_{n}) = A\cap (\bigcup B_{n}) \in \lambda(S)$, and hence $\bigcup B_{n} \in \cC_{A}$. Thus, $\cC_{A}$ is a $\lambda$-system. Since $\lambda(S)$ is the smallest $\lambda$-system containing $S$, we have $\lambda(S) = \cC_{A}$, and hence for any $B,C \in S$, we have $B,C \in S$ implies $B\cap C = A\cap C \in S$. Thus, $\lambda(S)$ is a $\pi$-system.
    \end{enumerate}
\end{proof}


\begin{lemma}
    Let $S$ be a $\pi$-system, and let $\Prob$ and $\Q$ be two probability measures on $(\Sigma,\sigma(S))$ such that $\Prob(A) = \Q(A)$ for all $A \in S$. Then, $\Prob(A) = \Q(A)$ for all $A \in \sigma(S)$.
\end{lemma}
\begin{proof}
    Let $\cC = \{B \in \sigma(S) \mid \Prob(B) = \Q(B)\}$. Then, $S \subseteq \cC$, and we show that $\cC$ is a $\lambda$-system. We have $\Omega \in \cC$ since $\Prob(\Omega) = 1 = \Q(\Omega)$. If $B_{1},B_{2} \in \cC$ with $B_{1} \subseteq B_{2}$, then $\Prob(B_{2}\setminus B_{1}) = \Prob(B_{2}) - \Prob(B_{1}) = \Q(B_{2}) - \Q(B_{1}) = \Q(B_{2}\setminus B_{1})$, and hence $B_{2}\setminus B_{1} \in \cC$. Finally, if $B_{1},B_{2},\ldots$ are in $\cC$, then $\Prob(A) = \lim_{n \to \infty} \Prob(A_n)$, and since $\Q(A) = \lim_{n \to \infty} \Q(A_n)$, we have $\Prob(A) = \Q(A)$.
\end{proof}